---
title: "StableTraits"
author: "docmanny"
date: "2023-10-23"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r Libraries and functions, include = F}
knitr::opts_chunk$set(
  eval=T #, 
  #echo=F, message = F, warning = F, 
  # out.width="6in", fig.pos = "H", out.extra = "",
  # fig.retina = 2, dpi = 300, dev=c('cairo_pdf', 'svg', 'png','cairo_ps')
)

# if(basename(getwd()) != "paper_PLOS") setwd("paper_PLOS")

set.seed(1234)
library(data.table)
library(tidyverse)
library(magrittr)
library(viridis)
options(readr.num_columns = 0)
library(ape)
library(geiger)
library(nlme)
# library(treeio)
# library(tidytree)
library(ggtree)
library(ggimage)
library(ggpubr)
library(ggsci)
library(ggstance)
library(plotly)
library(ggplotify)
library(UpSetR)
library(ggrepel)
library(magick)
library(gghighlight)
library(broom.mixed)
library(purrr)
library(kableExtra)
library(stargazer)
library(ggnewscale)
library(ComplexHeatmap)
library(rms)
library(patchwork)
library(phytools)
library(geiger)
library(nlme)
library(evomap)
library(magrittr)
```

## Body Size

```{r BS Trees}
size_rates_tree <- read.tree('../data/stableTraits/UphamEtAl2019_harmonizednames_withSizeLifeLn_size_concat.rates_tree') %>% 
  as_tibble
size_sq_tree <- read.tree('../data/stableTraits/UphamEtAl2019_harmonizednames_withSizeLifeLn_size_concat.sqchange_tree') %>% 
  as_tibble
size_brlens <- readr::read_tsv('../data/stableTraits/UphamEtAl2019_harmonizednames_withSizeLifeLn_size_concat.brlens')
size_ancstates <- readr::read_tsv('../data/stableTraits/UphamEtAl2019_harmonizednames_withSizeLifeLn_size_concat.ancstates')
```

```{r}
size_ancstates
```


## Lifespan

```{r Lifespan Trees}
life_rates_tree <- read.tree('../data/stableTraits/UphamEtAl2019_harmonizednames_withSizeLifeLn_life_concat.rates_tree') %>% 
  as_tibble
life_sq_tree <- read.tree('../data/stableTraits/UphamEtAl2019_harmonizednames_withSizeLifeLn_life_concat.sqchange_tree') %>% 
  as_tibble
life_brlens <- read_tsv('../data/stableTraits/UphamEtAl2019_harmonizednames_withSizeLifeLn_life_concat.brlens')
life_ancstates <- readr::read_tsv('../data/stableTraits/UphamEtAl2019_harmonizednames_withSizeLifeLn_life_concat.ancstates')
```

```{r}
life_ancstates
```


## Summary
```{r}
ls.stats <- dplyr::bind_rows(
  size_ancstates %>% dplyr::filter(stringr::str_starts(Parameter, "stable_")) %>% dplyr::mutate(Trait='Mass'),
  life_ancstates %>% dplyr::filter(stringr::str_starts(Parameter, "stable_")) %>% dplyr::mutate(Trait='Lifespan')
)
```

## PANCOVA


```{r}

tr <- read.tree('../data/stableTraits/UphamEtAl2019_harmonizednames_withSizeLifeLn_size_concat.tree')
# ggtree(size_sq_tree %>% as.treedata)
lifesize.dat <- dplyr::full_join(
  life_ancstates %>% dplyr::select(Parameter, lnLife=Median),
  size_ancstates %>% dplyr::select(Parameter, lnSize=Median)
) %>% 
  dplyr::filter(stringr::str_starts(Parameter, 'stable_', negate=T)) %>% 
  dplyr::rename(label='Parameter')
lifesize.trdat <- tidytree::full_join(tr, lifesize.dat)
# lifesize.trdat %>% ggtree(aes(color=lnLife))

lifesize.dat.noNode <- lifesize.dat %>% 
  dplyr::filter(label %in% tr$tip.label) %>% 
  tibble::column_to_rownames('label') %>% 
  as.data.frame
```

```{r}
data<-lifesize.dat.noNode
tree<-tr
Y<-"lnLife"
X<-"lnSize"
data<-data[,c(which(colnames(data)==Y),which(colnames(data)==X)),drop=F]
data<-na.omit(data)
tree<-geiger::treedata(tree,data,sort=T,warnings=F)$phy   #match the tree to the data
data<-as.data.frame(geiger::treedata(tree,data,sort=T,warnings=F)$data)   #match the data to the tree
colnames(data)<-c("Dependent","Independent")     #Note that naming the variables 'Dependent' and 'Independent' serves only to standardize the procedure below across different data sets.
```


```{r}
node.chiroptera <- tidytree::MRCA(tree, c('Myotis_lucifugus','Pteropus_vampyrus'))
node.laurasiatheria <- tidytree::MRCA(tree, c('Myotis_lucifugus','Sorex_fumeus'))
# tipnodes.chiroptera = tidytree::offspring(
#   lifesize.trdat, 
#   .node=node.chiroptera, 
#   tiponly = T
# )
# Chiroptera <- lifesize.trdat %>% 
#   tidytree::as_tibble() %>% 
#   dplyr::filter(node %in% tipnodes.chiroptera) %>% 
#   dplyr::pull(label)

Chiroptera <- getTips(tree, node.chiroptera)

NotChiroptera <- setdiff(getTips(tree, tidytree::rootnode(tr)),Chiroptera)

Laurasiatheria <- getTips(tree, node.laurasiatheria)

NotLaurasiatheria <- setdiff(getTips(tree, tidytree::rootnode(tree)),Laurasiatheria)


Laurasiatheria.NotChiroptera <- setdiff(getTips(tree, node.laurasiatheria), Chiroptera)

```

```{r plot all mammals, fig.height = 4, warnings = F, fig.width = 6, fig.align = "center"}
plot(
  lifesize.dat.noNode$lnLife~lifesize.dat.noNode$lnSize,
  col="white",
  pch=21,
  xlab="lnLife", 
  ylab="lnSize",
  asp=1,
  cex.lab=2
) 
pGLS.plotGrade('lnLife','lnSize',tree = tr %>% as.phylo(), data=lifesize.dat.noNode)


```

```{r plot bats vs others, fig.height = 4, warnings = F, fig.width = 6, fig.align = "center"}
plot(
  lifesize.dat.noNode$lnLife~lifesize.dat.noNode$lnSize,
  col="white",
  pch=21,
  xlab="lnLife", 
  ylab="lnSize",
  asp=1,
  cex.lab=2
) 
pGLS.plotGrade(
  'lnLife','lnSize',
  tree = tr %>% as.phylo(), 
  data=lifesize.dat.noNode,
  model='BM',
  group=NotChiroptera,
  col='grey',
  lwd=3, 
  cex=1,
  pch=21
)
pGLS.plotGrade(
  'lnLife','lnSize',
  tree = tr %>% as.phylo(), 
  data=lifesize.dat.noNode,
  model='BM',
  group=Laurasiatheria,
  col='violet',
  lwd=3, 
  cex=1,
  pch=21
)
pGLS.plotGrade(
  'lnLife','lnSize',
  tree = tr %>% as.phylo(), 
  data=lifesize.dat.noNode,
  model='BM',
  group=Chiroptera,
  col='green',
  lwd=3, 
  cex=1,
  pch=21
)
legend(15, 2, legend=c("Mammals", "Laurasiatheria", 'Bats'),
       fill=c("grey",'violet', "green"), cex=0.8)
```

### pANCOVA: Bats

```{r groups: chiroptera vs not-chiroptera}
#For differences in slope: 
grpS<-rep("A",length(rownames(data))) 
grpS[Chiroptera]<-"B" 
grpS<-as.factor(grpS) 
names(grpS)<-rownames(data)

#For differences in intercept: 
grpI<-rep("A",length(rownames(data))) 
grpI[Chiroptera]<-"B" 
grpI<-as.factor(grpI) 
names(grpI)<-rownames(data)
```

```{r Models for Bats}

Model<-model.matrix(as.formula(Dependent~Independent),data)

#(1) Differences in slopes, holding intercept constant: 
Model_S<-model.matrix(as.formula(Dependent~grpS:Independent),data) 

#(2) Differences in intercept, holding slopes constant: 
Model_I<-model.matrix(as.formula(Dependent~grpI + Independent),data) 

#(3) Differences in slopes and differences in intercept: 
Model_SI<-model.matrix(as.formula(Dependent~grpI + grpS:Independent),data)
```

```{r PANCOVA for Bats}
#(1) Differences in slopes, holding intercept constant:
gls.ancova(Dependent~Independent,vcv(tree),Model,Model_S)

#(2) Differences in intercept, holding slopes constant:
gls.ancova(Dependent~Independent,vcv(tree),Model,Model_I)

#(3) Differences in slopes and differences in intercept:
gls.ancova(Dependent~Independent,vcv(tree),Model,Model_SI)

#(4) Differences in slopes vs differences in both
gls.ancova(Dependent~Independent,vcv(tree),Model_I,Model_SI)
```

### pANCOVA: Laurasiatheria

```{r groups: Laurasiatheria vs not-Laurasiatheria}
#For differences in slope: 
grpS<-rep("A",length(rownames(data))) 
grpS[Laurasiatheria]<-"B" 
grpS<-as.factor(grpS) 
names(grpS)<-rownames(data)

#For differences in intercept: 
grpI<-rep("A",length(rownames(data))) 
grpI[Laurasiatheria]<-"B" 
grpI<-as.factor(grpI) 
names(grpI)<-rownames(data)
```

```{r Models for Laurasiatheria}

Model<-model.matrix(as.formula(Dependent~Independent),data)

#(1) Differences in slopes, holding intercept constant: 
Model_S<-model.matrix(as.formula(Dependent~grpS:Independent),data) 

#(2) Differences in intercept, holding slopes constant: 
Model_I<-model.matrix(as.formula(Dependent~grpI + Independent),data) 

#(3) Differences in slopes and differences in intercept: 
Model_SI<-model.matrix(as.formula(Dependent~grpI + grpS:Independent),data)
```

```{r PANCOVA for Laurasiatheria}
#(1) Differences in slopes, holding intercept constant:
gls.ancova(Dependent~Independent,vcv(tree),Model,Model_S)

#(2) Differences in intercept, holding slopes constant:
gls.ancova(Dependent~Independent,vcv(tree),Model,Model_I)

#(3) Differences in slopes and differences in intercept:
gls.ancova(Dependent~Independent,vcv(tree),Model,Model_SI)

#(4) Differences in slopes vs differences in both
gls.ancova(Dependent~Independent,vcv(tree),Model_I,Model_SI)
```

### pANCOVA: Laurasiatheria, no bats

```{r groups: Laurasiatheria.NotChiroptera vs not-Laurasiatheria}
#For differences in slope: 
grpS<-rep("A",length(rownames(data))) 
grpS[Laurasiatheria.NotChiroptera]<-"B" 
grpS<-as.factor(grpS) 
names(grpS)<-rownames(data)

#For differences in intercept: 
grpI<-rep("A",length(rownames(data))) 
grpI[Laurasiatheria.NotChiroptera]<-"B" 
grpI<-as.factor(grpI) 
names(grpI)<-rownames(data)
```

```{r Models for Laurasiatheria.NotChiroptera}

Model<-model.matrix(as.formula(Dependent~Independent),data)

#(1) Differences in slopes, holding intercept constant: 
Model_S<-model.matrix(as.formula(Dependent~grpS:Independent),data) 

#(2) Differences in intercept, holding slopes constant: 
Model_I<-model.matrix(as.formula(Dependent~grpI + Independent),data) 

#(3) Differences in slopes and differences in intercept: 
Model_SI<-model.matrix(as.formula(Dependent~grpI + grpS:Independent),data)
```

```{r PANCOVA for Laurasiatheria.NotChiropteraa}
#(1) Differences in slopes, holding intercept constant:
gls.ancova(Dependent~Independent,vcv(tree),Model,Model_S)

#(2) Differences in intercept, holding slopes constant:
gls.ancova(Dependent~Independent,vcv(tree),Model,Model_I)

#(3) Differences in slopes and differences in intercept:
gls.ancova(Dependent~Independent,vcv(tree),Model,Model_SI)

#(4) Differences in slopes vs differences in both
gls.ancova(Dependent~Independent,vcv(tree),Model_I,Model_SI)
```

### pANCOVA: Yangochiroptera vs Yinpterochiroptera

```{r}
tree.og <- tree
data.og <- data
tree<-tidytree::tree_subset(tree, node = tidytree::MRCA(tree, c('Myotis_lucifugus','Pteropus_vampyrus')), levels_back = 0)

tree<-geiger::treedata(tree,data,sort=T,warnings=F)$phy   #match the tree to the data
data<-as.data.frame(geiger::treedata(tree,data,sort=T,warnings=F)$data)   #match the data to the tree
colnames(data)<-c("Dependent","Independent")     #Note that naming the variables 'Dependent' and 'Independent' serves only to standardize the procedure below across different data sets.
```


```{r}

node.yinpterochiroptera = tidytree::MRCA(tree, 'Macroderma_gigas', 'Pteropus_giganteus')
node.yangochiroptera = tidytree::MRCA(tree, 'Myotis_lucifugus', 'Mystacina_tuberculata') 
# tipnodes.chiroptera = tidytree::offspring(
#   lifesize.trdat, 
#   .node=node.chiroptera, 
#   tiponly = T
# )
# Chiroptera <- lifesize.trdat %>% 
#   tidytree::as_tibble() %>% 
#   dplyr::filter(node %in% tipnodes.chiroptera) %>% 
#   dplyr::pull(label)

Yinpterochiroptera <- getTips(tree, node.yinpterochiroptera)

Yangochiroptera <- getTips(tree,node.yangochiroptera)

```

```{r, fig.height = 4, warnings = F, fig.width = 6, fig.align = "center"}
plot(data$Dependent~data$Independent,col="white",pch=21,xlab="", ylab="",asp=1,cex.lab=2) 
pGLS.plotGrade('Dependent','Independent',tree = tree, data=data)
```

```{r plot yinptero vs yango, fig.height = 4, warnings = F, fig.width = 6, fig.align = "center"}
plot(data$Dependent~data$Independent,col="white",pch=21,xlab="", ylab="",asp=1,cex.lab=2) 
pGLS.plotGrade(
  'Dependent','Independent',
  tree = tree, data=data,
  model='BM',
  group=Yangochiroptera,
  col='violet',
  lwd=3, 
  cex=1,
  pch=21
)
pGLS.plotGrade(
  'Dependent','Independent',
  tree = tree, data=data,
  model='BM',
  group=Yinpterochiroptera,
  col='green',
  lwd=3, 
  cex=1,
  pch=21
)

legend(6, 2, legend=c("Yangochiroptera", 'Yinpterochiroptera'),
       fill=c('violet', "green"), cex=0.8)
```

```{r groups: Yangochiroptera vs Yinpterochiroptera}
#For differences in slope: 
grpS<-rep("A",length(rownames(data))) 
grpS[Yangochiroptera]<-"B" 
grpS<-as.factor(grpS) 
names(grpS)<-rownames(data)

#For differences in intercept: 
grpI<-rep("A",length(rownames(data))) 
grpI[Yangochiroptera]<-"B" 
grpI<-as.factor(grpI) 
names(grpI)<-rownames(data)
```

```{r Models for Yangochiroptera}

Model<-model.matrix(as.formula(Dependent~Independent),data)

#(1) Differences in slopes, holding intercept constant: 
Model_S<-model.matrix(as.formula(Dependent~grpS:Independent),data) 

#(2) Differences in intercept, holding slopes constant: 
Model_I<-model.matrix(as.formula(Dependent~grpI + Independent),data) 

#(3) Differences in slopes and differences in intercept: 
Model_SI<-model.matrix(as.formula(Dependent~grpI + grpS:Independent),data)
```

```{r PANCOVA for Yangochiroptera}
#(1) Differences in slopes, holding intercept constant:
gls.ancova(Dependent~Independent,vcv(tree),Model,Model_S)

#(2) Differences in intercept, holding slopes constant:
gls.ancova(Dependent~Independent,vcv(tree),Model,Model_I)

#(3) Differences in slopes and differences in intercept:
gls.ancova(Dependent~Independent,vcv(tree),Model,Model_SI)

#(4) Differences in slopes vs differences in both
gls.ancova(Dependent~Independent,vcv(tree),Model_I,Model_SI)
```


### pANCOVA: Within Yangochiroptera

```{r}
node.vespertilionidae = tidytree::MRCA(tree, 'Myotis_lucifugus', 'Lasiurus_cinereus')
node.myotis = tidytree::MRCA(tree, 'Myotis_lucifugus', 'Myotis_emarginatus')
node.molossidae = tidytree::MRCA(tree, 'Molossus_molossus', 'Chaerephon_pumilus')
node.phyllostomidae = tidytree::MRCA(tree, 'Macrotus_waterhousii', 'Artibeus_jamaicensis')

Vespertilionidae <- getTips(tree, node.vespertilionidae)
Myotis <- getTips(tree, node.myotis)
Molossidae <- getTips(tree, node.molossidae)
Phyllostomidae <- getTips(tree, node.phyllostomidae)

Yangochiroptera.Other <- setdiff(
  Yangochiroptera,
  Vespertilionidae
) %>% 
  setdiff(
    .,
    Molossidae
  ) %>% 
  setdiff(
    .,
    Phyllostomidae
  )
Vespertilionidae.NotMyotis <- setdiff(
  Vespertilionidae,
  Myotis
)
```


```{r plot bat subtypes, fig.height = 4, warnings = F, fig.width = 6, fig.align = "center"}
plot(data$Dependent~data$Independent,col="white",pch=21,xlab="", ylab="",asp=1,cex.lab=2) 
pGLS.plotGrade(
  'Dependent','Independent',
  tree = tree, data=data,
  model='BM',
  group=Yinpterochiroptera,
  col='black',
  lwd=3, 
  cex=1,
  pch=21
)
pGLS.plotGrade(
  'Dependent','Independent',
  tree = tree, data=data,
  model='BM',
  group=Yangochiroptera.Other,
  col='violet',
  lwd=3, 
  cex=1,
  pch=21
)
pGLS.plotGrade(
  'Dependent','Independent',
  tree = tree, data=data,
  model='BM',
  group=Vespertilionidae.NotMyotis,
  col='green',
  lwd=3, 
  cex=1,
  pch=21
)
pGLS.plotGrade(
  'Dependent','Independent',
  tree = tree, data=data,
  model='BM',
  group=Myotis,
  col='blue',
  lwd=3, 
  cex=1,
  pch=21
)
pGLS.plotGrade(
  'Dependent','Independent',
  tree = tree, data=data,
  model='BM',
  group=Molossidae,
  col='yellow',
  lwd=3, 
  cex=1,
  pch=21
)
pGLS.plotGrade(
  'Dependent','Independent',
  tree = tree, data=data,
  model='BM',
  group=Phyllostomidae,
  col='red',
  lwd=3, 
  cex=1,
  pch=21
)

legend(6.25, 2.5,
        legend=c(
          'Yinpterochiroptera',
          'Yangochiroptera.Other',
          'Vespertilionidae.NotMyotis',
          'Myotis',
          'Molossidae',
          'Phyllostomidae'
        ),
       fill=c(
          'black',
          'violet',
          'green',
          'blue',
          'yellow',
          'red'
       ), 
       cex=0.8)
```

```{r groups: allGroups}
#For differences in slope: 
grpS<-rep("A",length(rownames(data))) 
grpS[Vespertilionidae]<-"B" 
grpS[Myotis]<-"C"
grpS[Molossidae]<-"D"
grpS[Phyllostomidae]<-"E"  
grpS<-as.factor(grpS) 
names(grpS)<-rownames(data)

#For differences in intercept: 
grpI<-rep("A",length(rownames(data))) 
grpI[Vespertilionidae]<-"B" 
grpI[Myotis]<-"C"
grpI[Molossidae]<-"D"
grpI[Phyllostomidae]<-"E" 
grpI<-as.factor(grpI) 
names(grpI)<-rownames(data)
```

```{r Models for allGroups}

Model<-model.matrix(as.formula(Dependent~Independent),data)

#(1) Differences in slopes, holding intercept constant: 
Model_S<-model.matrix(as.formula(Dependent~grpS:Independent),data) 

#(2) Differences in intercept, holding slopes constant: 
Model_I<-model.matrix(as.formula(Dependent~grpI + Independent),data) 

#(3) Differences in slopes and differences in intercept: 
Model_SI<-model.matrix(as.formula(Dependent~grpI + grpS:Independent),data)
```

```{r PANCOVA for allGroups}
#(1) Differences in slopes, holding intercept constant:
gls.ancova(Dependent~Independent,vcv(tree),Model,Model_S)

#(2) Differences in intercept, holding slopes constant:
gls.ancova(Dependent~Independent,vcv(tree),Model,Model_I)

#(3) Differences in slopes and differences in intercept:
gls.ancova(Dependent~Independent,vcv(tree),Model,Model_SI)

#(4) Differences in slopes vs differences in both
gls.ancova(Dependent~Independent,vcv(tree),Model_I,Model_SI)
```

### PANCOVA: drop Myoits auriculus

```{r}
tree <- tree.og
data <- data.og
tree <- tree %>% tidytree::drop.tip(c('Myotis_auriculus'))
data <- data[which(rownames(data) != "Myotis_auriculus"),]
node.vespertilionidae = tidytree::MRCA(tree, 'Myotis_lucifugus', 'Lasiurus_cinereus')
node.myotis = tidytree::MRCA(tree, 'Myotis_lucifugus', 'Myotis_emarginatus')
node.molossidae = tidytree::MRCA(tree, 'Molossus_molossus', 'Chaerephon_pumilus')
node.phyllostomidae = tidytree::MRCA(tree, 'Macrotus_waterhousii', 'Artibeus_jamaicensis')

node.chiroptera <- tidytree::MRCA(tree, c('Myotis_lucifugus','Pteropus_vampyrus'))
node.laurasiatheria <- tidytree::MRCA(tree, c('Myotis_lucifugus','Sorex_fumeus'))
Chiroptera <- getTips(tree, node.chiroptera)

NotChiroptera <- setdiff(getTips(tree, tidytree::rootnode(tree)),Chiroptera)

Laurasiatheria <- getTips(tree, node.laurasiatheria)

NotLaurasiatheria <- setdiff(getTips(tree, tidytree::rootnode(tree)),Laurasiatheria)


Laurasiatheria.NotChiroptera <- setdiff(getTips(tree, node.laurasiatheria), Chiroptera)
Vespertilionidae <- getTips(tree, node.vespertilionidae)
Myotis <- getTips(tree, node.myotis)
Molossidae <- getTips(tree, node.molossidae)
Phyllostomidae <- getTips(tree, node.phyllostomidae)

Yangochiroptera.Other <- setdiff(
  Yangochiroptera,
  Vespertilionidae
) %>% 
  setdiff(
    .,
    Molossidae
  ) %>% 
  setdiff(
    .,
    Phyllostomidae
  )
Vespertilionidae.NotMyotis <- setdiff(
  Vespertilionidae,
  Myotis
)
```

```{r plot bats vs others no myoAui, fig.height = 4, warnings = F, fig.width = 6, fig.align = "center"}
plot(
  data$Dependent~data$Independent,
  col="white",
  pch=21,
  xlab="Independent", 
  ylab="Dependent",
  asp=1,
  cex.lab=2
) 
pGLS.plotGrade(
  'Independent','Dependent',
  tree = tree, 
  data= data,
  model='BM',
  group=NotChiroptera,
  col='grey',
  lwd=3, 
  cex=1,
  pch=21
)
pGLS.plotGrade(
  'Independent','Dependent',
  tree = tree, 
  data= data,
  model='BM',
  group=Laurasiatheria,
  col='violet',
  lwd=3, 
  cex=1,
  pch=21
)
pGLS.plotGrade(
  'Independent','Dependent',
  tree = tree, 
  data= data,
  model='BM',
  group=Chiroptera,
  col='green',
  lwd=3, 
  cex=1,
  pch=21
)
legend(15, 2, legend=c("Mammals", "Laurasiatheria", 'Bats'),
       fill=c("grey",'violet', "green"), cex=0.8)
```

```{r plot yinptero vs yango no myoAui, fig.height = 4, warnings = F, fig.width = 6, fig.align = "center"}
plot(
  data$Dependent~data$Independent,
  col="white",
  pch=21,
  xlab="Dependent", 
  ylab="Independent",
  asp=1,
  cex.lab=2
) 
pGLS.plotGrade(
  'Dependent','Independent',
  tree = tree, 
  data= data,
  model='BM',
  group=NotChiroptera,
  col='grey',
  lwd=3, 
  cex=1,
  pch=21
)
pGLS.plotGrade(
  'Dependent','Independent',
  tree = tree, data=data,
  model='BM',
  group=Yangochiroptera,
  col='violet',
  lwd=3, 
  cex=1,
  pch=21
)
pGLS.plotGrade(
  'Dependent','Independent',
  tree = tree, data=data,
  model='BM',
  group=Yinpterochiroptera,
  col='green',
  lwd=3, 
  cex=1,
  pch=21
)

legend(6, 2, legend=c("Other Mammals", "Yangochiroptera", 'Yinpterochiroptera'),
       fill=c("grey",'violet', "green"), cex=0.8)
```

```{r}
tree<-tidytree::tree_subset(tree, node = tidytree::MRCA(tree, c('Myotis_lucifugus','Pteropus_vampyrus')), levels_back = 0)

tree<-geiger::treedata(tree,data,sort=T,warnings=F)$phy   #match the tree to the data
data<-as.data.frame(geiger::treedata(tree,data,sort=T,warnings=F)$data)   #match the data to the tree
colnames(data)<-c("Dependent","Independent")     #Note that naming the variables 'Dependent' and 'Independent' serves only to standardize the procedure below across different data sets.
node.yinpterochiroptera = tidytree::MRCA(tree, 'Macroderma_gigas', 'Pteropus_giganteus')
node.yangochiroptera = tidytree::MRCA(tree, 'Myotis_lucifugus', 'Mystacina_tuberculata') 
# tipnodes.chiroptera = tidytree::offspring(
#   lifesize.trdat, 
#   .node=node.chiroptera, 
#   tiponly = T
# )
# Chiroptera <- lifesize.trdat %>% 
#   tidytree::as_tibble() %>% 
#   dplyr::filter(node %in% tipnodes.chiroptera) %>% 
#   dplyr::pull(label)

Yinpterochiroptera <- getTips(tree, node.yinpterochiroptera)

Yangochiroptera <- getTips(tree,node.yangochiroptera)
node.vespertilionidae = tidytree::MRCA(tree, 'Myotis_lucifugus', 'Lasiurus_cinereus')
node.myotis = tidytree::MRCA(tree, 'Myotis_lucifugus', 'Myotis_emarginatus')
node.molossidae = tidytree::MRCA(tree, 'Molossus_molossus', 'Chaerephon_pumilus')
node.phyllostomidae = tidytree::MRCA(tree, 'Macrotus_waterhousii', 'Artibeus_jamaicensis')

Vespertilionidae <- getTips(tree, node.vespertilionidae)
Myotis <- getTips(tree, node.myotis)
Molossidae <- getTips(tree, node.molossidae)
Phyllostomidae <- getTips(tree, node.phyllostomidae)

Yangochiroptera.Other <- setdiff(
  Yangochiroptera,
  Vespertilionidae
) %>% 
  setdiff(
    .,
    Molossidae
  ) %>% 
  setdiff(
    .,
    Phyllostomidae
  )
Vespertilionidae.NotMyotis <- setdiff(
  Vespertilionidae,
  Myotis
)
```

```{r plot bat subtypes without auriculus, fig.height = 4, warnings = F, fig.width = 6, fig.align = "center"}
plot(data$Dependent~data$Independent,col="white",pch=21,xlab="", ylab="",asp=1,cex.lab=2) 
pGLS.plotGrade(
  'Dependent','Independent',
  tree = tree, data=data,
  model='BM',
  group=Yinpterochiroptera,
  col='black',
  lwd=3, 
  cex=1,
  pch=21
)
pGLS.plotGrade(
  'Dependent','Independent',
  tree = tree, data=data,
  model='BM',
  group=Yangochiroptera.Other,
  col='violet',
  lwd=3, 
  cex=1,
  pch=21
)
pGLS.plotGrade(
  'Dependent','Independent',
  tree = tree, data=data,
  model='BM',
  group=Vespertilionidae.NotMyotis,
  col='green',
  lwd=3, 
  cex=1,
  pch=21
)
pGLS.plotGrade(
  'Dependent','Independent',
  tree = tree, data=data,
  model='BM',
  group=Myotis,
  col='blue',
  lwd=3, 
  cex=1,
  pch=21
)
pGLS.plotGrade(
  'Dependent','Independent',
  tree = tree, data=data,
  model='BM',
  group=Molossidae,
  col='yellow',
  lwd=3, 
  cex=1,
  pch=21
)
pGLS.plotGrade(
  'Dependent','Independent',
  tree = tree, data=data,
  model='BM',
  group=Phyllostomidae,
  col='red',
  lwd=3, 
  cex=1,
  pch=21
)
legend(6.25, 2.5,
        legend=c(
          'Yinpterochiroptera',
          'Yangochiroptera.Other',
          'Vespertilionidae.NotMyotis',
          'Myotis',
          'Molossidae',
          'Phyllostomidae'
        ),
       fill=c(
          'black',
          'violet',
          'green',
          'blue',
          'yellow',
          'red'
       ), 
       cex=0.8)
```

```{r groups: allGroups no myoAui}
#For differences in slope: 
grpS<-rep("A",length(rownames(data))) 
grpS[Vespertilionidae]<-"B" 
grpS[Myotis]<-"C"
grpS[Molossidae]<-"D"
grpS[Phyllostomidae]<-"E"  
grpS<-as.factor(grpS) 
names(grpS)<-rownames(data)

#For differences in intercept: 
grpI<-rep("A",length(rownames(data))) 
grpI[Vespertilionidae]<-"B" 
grpI[Myotis]<-"C"
grpI[Molossidae]<-"D"
grpI[Phyllostomidae]<-"E" 
grpI<-as.factor(grpI) 
names(grpI)<-rownames(data)
```

```{r Models for allGroups no myoAui}

Model<-model.matrix(as.formula(Dependent~Independent),data)

#(1) Differences in slopes, holding intercept constant: 
Model_S<-model.matrix(as.formula(Dependent~grpS:Independent),data) 

#(2) Differences in intercept, holding slopes constant: 
Model_I<-model.matrix(as.formula(Dependent~grpI + Independent),data) 

#(3) Differences in slopes and differences in intercept: 
Model_SI<-model.matrix(as.formula(Dependent~grpI + grpS:Independent),data)
```

```{r PANCOVA for allGroups no myoAui}
#(1) Differences in slopes, holding intercept constant:
gls.ancova(Dependent~Independent,vcv(tree),Model,Model_S)

#(2) Differences in intercept, holding slopes constant:
gls.ancova(Dependent~Independent,vcv(tree),Model,Model_I)

#(3) Differences in slopes and differences in intercept:
gls.ancova(Dependent~Independent,vcv(tree),Model,Model_SI)

#(4) Differences in slopes vs differences in both
gls.ancova(Dependent~Independent,vcv(tree),Model_I,Model_SI)
```

### Remarks
In conclusion, bats are weird. However, its not all bats, but really its just Yangochiroptera. In fact, within Yangochiroptera, most families either follow the mold (like Molossidae) or have no correlation (like Vespertilionidae and Phyllostomidae). Myotis auriculus is definitely an outlier - with it included, there is a negative correlation between body size and lifespan, but without it, genus Myotis follows the same trends as Vespertilionidae in general.


